---
title: "CT Active Licensed Hospice and Home Healthcare Facilitites"
author: 
  - name: Alexander Senetcky
    id: as
    orcid: 0009-0009-3730-5397
    email: alexander.senetcky@ct.gov
    affiliation:
      - name: Connecticut Department of Public Health
        city: Hartford
        state: CT
        url: https://portal.ct.gov/dph
format: 
  html:
    code-fold: true
    embed-resources: true
    toc: true
    toc-location: left
    other-links:
      - text: Connecticut Open Data Portal
        href: https://data.ct.gov/
      - text: State Licenses and Credentials
        href: https://data.ct.gov/Business/State-Licenses-and-Credentials/ngch-56tr/about_data
      - text: Facility Credential Types
        href: https://data.ct.gov/Business/Credential-Types/rykc-pttg/about_data
date: "2025-01-29"
---

## Project Overview

The goal of this project was to map Home Health Care and Hospice Licensed
facilities in Connecticut using open data from the 
[Connecticut Open Data Portal](https://data.ct.gov/) (ODP).

The State License and Credential dataset is large, unordered and updated daily.
The ODP API backend was used to pull down all active Home Health Care and
Hospice facilities and ignore the other million or so rows. There appear to be
93 faciliities in the lastest parts of the data, with 90 being "Active" or
"Active in renewal".  Of those active 90, only 2 were credentialed as hospice
facilities.

In the short amount of time given I was able to grab the addresses from these 90
facilities and use the Census geocoder API to get the longitutude and latitude
of each facility. 87 of these facilities were able to be geocoded and
unfortunately one of those was a hospice facility. I am confident that in an
intentional production environment we could get 100% of those geocoded.

There may also be some facilities licensed as Home Health Care that also
serve as a hospice.



```{r}
#| label: "setup"
#| include: false

library(dplyr)
library(lubridate)
library(tigris)
library(RSocrata)
library(stringr)
library(glue)
library(tigris)
library(mapview)
library(tidygeocoder)
library(sf)
library(gt)

```

## Facilities

```{r}
#| label: "grab-data"

# splitting this out in case iteration is needed

domain <- "https://data.ct.gov/"
resource <- "/resource/ngch-56tr.json"
where_clause <- "?$where="
expiration_clause <- "expirationdate >= '2025-01-28'"
active_clause <- "active = 1"
credential_type_clause <- "credentialtype in('HHC', 'HSPC')"

# grab all facilities that haven't expired as of 2025-01-28
include_non_active <- glue(
  "{domain}{resource}{where_clause}{expiration_clause} and {credential_type_clause}"
)

# grab all active facilities
only_active <-  glue(
  "{domain}{resource}{where_clause}{active_clause} and {credential_type_clause}"
)


```

