---
title: "CT Active Licensed Hospice and Home Healthcare Facilitites"
author: 
  - name: Alexander Senetcky
    id: as
    orcid: 0009-0009-3730-5397
    email: alexander.senetcky@ct.gov
    affiliation:
      - name: Connecticut Department of Public Health
        city: Hartford
        state: CT
        url: https://portal.ct.gov/dph
format: 
  html:
    code-fold: true
    embed-resources: true
    toc: true
    toc-location: left
    other-links:
      - text: Connecticut Open Data Portal
        href: https://data.ct.gov/
      - text: State Licenses and Credentials
        href: https://data.ct.gov/Business/State-Licenses-and-Credentials/ngch-56tr/about_data
      - text: Facility Credential Types
        href: https://data.ct.gov/Business/Credential-Types/rykc-pttg/about_data
      - text: Census Geocoder
        href: https://geocoding.geo.census.gov/geocoder/
      - text: Census TIGER/Line Shapefiles
        href: https://www.census.gov/geographies/mapping-files/time-series/geo/tiger-line-file.html
date: "2025-01-29"
---

## Project Overview

The goal of this project was to map Home Health Care and Hospice Licensed
facilities in Connecticut using open data from the 
[Connecticut Open Data Portal](https://data.ct.gov/) (ODP).

The State License and Credential dataset is large, unordered and updated daily.
The ODP API backend was used to pull down all active Home Health Care and
Hospice facilities and ignore the other million or so rows. There appear to be
93 facilities in the lastest parts of the data, with 90 being "Active" or
"Active in renewal".  Of those active 90, only 2 were credentialed as hospice
facilities.

In the short amount of time given I was able to grab the addresses from these 90
facilities and use the Census geocoder API to get the longitutude and latitude
of each facility. 87 of these facilities were able to be geocoded and
unfortunately one of those was a hospice facility. I am confident that in an
intentional production environment we could get 100% of those geocoded.

There may also be some facilities licensed as Home Health Care that also
serve as a hospice.

For a list of active and inactive facilities, please see @fig-facility-table.

Please reference @fig-facility-map to see the geocoded results displayed on
a map.



```{r}
#| label: "setup"
#| include: false

# Libraries
library(dplyr)
library(lubridate)
library(tigris)
library(RSocrata)
library(stringr)
library(glue)
library(tigris)
library(mapview)
library(tidygeocoder)
library(sf)
library(DT)
library(readr)
library(fs)

#options
options(tigris_use_cache = TRUE)

```

## The Process

### API Data Request

First the data is pulled from the API on the ODP.  Only facilities
that are credentialed as "HSSPC" and "HHC" are included. To get at
anything that isn't too old but may not be active, the `expirationdate` 
variable is used and any facility that hasn't expired as of 2025-01-28
is included. 

::: {.callout-note}
Note that these is *no* metadata about the contents of the credential data
provided. So the use of `expirationdate` was a best guess.  If a wider net
needs to be cast, other combinations of variables and values can be 
used.
:::

```{r}
#| label: "grab-data"

# splitting this out in case iteration is needed
domain <- "https://data.ct.gov/"
resource <- "resource/ngch-56tr.json"
where_clause <- "?$where="
expiration_clause <- "expirationdate >= '2025-01-28'"
active_clause <- "active = 1"
credential_type_clause <- "credentialtype in('HHC', 'HSPC')"

# grab all facilities that haven't expired as of 2025-01-28
include_non_active_statement <- glue(
  "{domain}{resource}{where_clause}{expiration_clause} and {credential_type_clause}"
)

# grab all active facilities
only_active_statement <-  glue(
  "{domain}{resource}{where_clause}{active_clause} and {credential_type_clause}"
)

include_non_active <- 
  RSocrata::read.socrata(url = include_non_active_statement) |> 
  as_tibble()

only_active <- 
  RSocrata::read.socrata(only_active_statement) |> 
  as_tibble()
```

The raw data look like this:

```{r}
glimpse(include_non_active)
```


### Clean and Geocode

Next up the data are cleaned and the relevant elements are kept.

```{r}
#| label: clean-data

# At the time of writing the two datasets are the same - minus the 3 non-active
# facilities, so let's use that as our data layer

facilities <- 
  include_non_active |> 
  select(
    name,
    type,
    fullcredentialcode,
    credentialnumber,
    credential,
    status,
    issuedate,
    effectivedate,
    expirationdate,
    address,
    city,
    state,
    zip,
    recordrefreshedon
  ) |>
  mutate(
    across(
      where(lubridate::is.timepoint),
      lubridate::as_date
    ),
    zip = str_sub(zip, end = -5L),
    across(
      c(name, type, status), str_to_title
    ),
    zip = if_else(zip == 0, NA_character_, zip),
    full_address = glue(
      "{address}, {city}, {state}"
    ),
    full_address = if_else(
      !is.na(zip),
      glue("{full_address} {zip}"),
      full_address
    )
  ) |>
  select(
    -c(
      address,
      city,
      state,
      zip
    )
  ) |> 
  dplyr::relocate(
    full_address,
    .after = expirationdate
  ) |> 
  dplyr::relocate(
    recordrefreshedon,
    .after = last_col()
  ) |> 
  arrange(status, credential, name)

# write out as .csv
# write_csv(
#   facilities,
#   file = fs::path_wd("data/hospice_hhc_facility_list.csv")
#   na = ""
# )

```

Then these data are run through the [Census Geocoder](https://geocoding.geo.census.gov/geocoder/).

```{r}
#| label: geocode-data
#| message: false

# geocode with census geocoder
# then drop rows not geocoded
geocoded_facilities <- 
  facilities |> 
  filter(status != "Closed") |> 
  tidygeocoder::geocode(
     address = full_address, method = "census"
  ) |> 
  filter(!is.na(long))

# convert to sf
geocoded_facilities <- 
  sf::st_as_sf(
    geocoded_facilities,
    coords = c("long", "lat"),
    crs = "NAD83"
  )
```

Shapefiles will also be needed as well. County Subdivision Shapefiles for
Connecticut will be grabbed from the
[Census TIGER/Line Shapefiles](https://www.census.gov/geographies/mapping-files/time-series/geo/tiger-line-file.html).

```{r}
#| label: shapefiles-town
#| message: false
#| warning: false

town_shapefile <- 
  tigris::county_subdivisions(
    state = "09",
    cb = TRUE,
    year = 2022,
    progress_bar = FALSE
  )

```


## Table

Below is table meant for a quick reference to the data.  Please 
see the provided .csv file for deeper analysis: `hospice_hhc_Facility_list.csv`

### Facility Info

```{r}
#| label: fig-facility-table

facilities |> 
  mutate(
    status = factor(status),
    credential = factor(credential),
    full_address = str_to_lower(full_address)
  ) |> 
  select(
    -c(
      type, 
      fullcredentialcode, 
      credentialnumber, 
      effectivedate,
      recordrefreshedon
    )
  ) |> 
  DT::datatable(
    filter = "top",
    rownames = FALSE,
    colnames = c(
      "Name",
      "Credential Type",
      "Status",
      "Issue Date",
      "Expiration Date",
      "Address"
    ),
    options = list(
      pageLength = 5,
      autoWidth = TRUE
    )
  )
```


## Map

### Active Connecticut Home Healthcare and Hospice Facilities

```{r}
#| label: fig-facility-map

map_data <- 
  geocoded_facilities |> 
     select(
      name,
      type,
      credential,
      fullcredentialcode,
      full_address,
      status,
      issuedate,
      effectivedate,
      expirationdate,
      geometry
    ) |> 
  rename(
    Name = name,
    `Business Type` = type,
    `Credential Type` = credential,
    Credential = fullcredentialcode,
    Address = full_address,
    Status = status,
    `Issue Date` = issuedate,
    `Effective Date` = effectivedate,
    `Expiration Date` = expirationdate
  )

mapview::mapview(
  town_shapefile,
  layer.name = "Town",
  alpha = 0.9,
  alpha.regions = 0.15,
  col.regions = "#3371E7",
  label = "NAME"
  #zcol = "NAME",
  #legend = TRUE,
  #burst = TRUE
) +
mapview::mapview(
  map_data,
  layer.name = "Facilities",
  # ycol = "lat",
  # xcol = "long",
  label = "Name",
  zcol = "Credential Type"
) 
  
```

